Trajectory Classification

### 2023-03-20 paper.qmd
trajectory classification
get some stuff to github

@@

some cool text.

Lorem Ipsum is simply dummy text of the printing and typesetting industry.
Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book.
It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged.
It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

@@

@@

## Open Data for Operational Performance Monitoring

@@

sequence separation

>>>>

Operational efficiency is a key element of addressing aviation's contribution to climate change.
<!-- verbatim copy --> Emissions may increase due to the expected growth in international air traffic until lower emitting technologies and fuels and other mitigating measures are developed and deployed ..
<!-- end copy --> ICAO adopted at its 41st Assembly a long-term aspirational goal for international aviation [@icao2022].
In support to the Paris Agreement, the goal is to achieve net-zero carbon emissions by 2050.

Levers for fuel reduction:

\* operational efficiency

\* market-based measures

\* sustainable aviation fuel

\* new aircraft propulsion and airframes

A wider use and pick-up of sustainable aviation fuel, and new aircraft propulsion technologies or aircraft design requires further research.

Despite the introduction of an initial market-based mechanism, immediate action to curb fuel burn and CO2 emissions rests with improvements of operational efficiency.

@@

@@

### Trajectory Flight Phase Segmentation & Milestone

Different approaches exists to detect and describe aircraft flight phases, e.g. recent machine learning algorithm [@sun2017flightphase].
This paper implements a heuristic approach with a focus on the detection of arrival traffic at the study airports.

### Landing Runway Identification

@@

@@

Spacing deviation

Let us consider a pair of consecutive landing aircraft denoted leader and trailer, with s their required time spacing1.
Using the constant time delay principle, the spacing deviation (or spacing error) at time t considers the current position of trailer at time t, and the past position of leader at time t -- s.
Precisely, it is defined as the difference between the respective minimum times from these two positions (see figure below): spacing deviation (t) = min time (trailer (t)) -- min time (leader (t -- s))

### 2023-03-20 x_trj-classification-segmentation.qmd
get some stuff to github

source(here::here("R", "level_segments_check.r"))

@@

ggplot() +
  geom_sf(data = rwys_egll_box) +
  geom_sf(data = baw262_ls)

```

@@

>>>>

source(here::here("R", "level_segments_check.R"))
source(here::here("R", "altitude_outlier_filter.R"))

@@

ggplot() +
  geom_sf(data = rwys_egll_box) +
  geom_sf(data = baw262_ls)
```

@@

## Start data preparation

work with dataset for a day

```{r}
source("./R/osn_dumps.R")
source("./R/osn_read_tardump.R")
source("./R/osn_make_nice_names.R")
source("./R/coerce_meter_to_feet.R")
```

```{r}
#| message: false
pth_to_dumps <- "../__DATA/OSN-dumps"
ptn <- "2019-08-01"

# list downloaded dumps
fns <- list.files(path = pth_to_dumps, pattern = ptn, full.names = TRUE)

# load list of chosen dumps
ds  <- fns |> 
  purrr::map_dfr(.f = read_osn_tar_file) |> 
  make_nice_names_osn() |> 
  mutate(ALT_B = coerce_meter_to_feet(ALT_B))
```

* filter pts within airport box (join per day runs for ~ 5mins)

```{r}
# EGLL airport box
box_sf  <- rwys_egll_box |> sf::st_as_sf() |> mutate(REF = "EGLL-box") |> rename(geometry = x)

# spatial subsetting, check alt, get ICAO24
# command x[y, ] is identical to x[y, , op = st_intersects]
seen_at_airport <- function(.pts, .poly_sf, .alt_threshold = 1000, .drop_coord = FALSE,...) {
  ds_pts  <- .pts |> cast_latlon_to_pts(.drop_coord = .drop_coord)
  # spatial subsetting, check alt, get ICAO24
  # command x[y, ] is identical to x[y, , op = st_intersects]
  inbox   <- ds_pts[.poly_sf,] |> 
    dplyr::filter(ALT_B < .alt_threshold)
  return(inbox)
}
```

```{r}
at_egll <- ds |> seen_at_airport(box_sf) 
at_egll <- at_egll |> mutate(SEEN = "AT EGLL") |> 
  summarise(BOX_MIN)
```

at egll with is.na(FLTID)
# A tibble: 214 × 1
   ICAO24
   <chr> 
 1 4951c7
 2 400941
 3 4ca13d
 4 4006b0
 5 3c65cd
 6 406a3e
 7 4b17fc
 8 406ae3
 9 4010dd
10 4009da

Load validation data set

```{}
egll_fact <- read_csv2("~/RProjects/__DATA/EGLL/EGLL_2019_FACT.csv")
egll_fact <- egll_fact |> 
  select(FLTID = AP_C_FLTID, REG = AP_C_REG, ADEP = ADEP_ICAO, ADES = ADES_ICAO
         ,MVT_TIME = MVT_TIME_UTC, BLOCK_TIME = BLOCK_TIME_UTC, PHASE = SRC_PHASE
         ,TYPE = ARCTYP, CLASS = AC_CLASS, RWY = AP_C_RWY, STND = AP_C_STND)
write_csv(egll_fact, "./data/validation-EGLL-2019.csv.gz")
```
```{r}
egll_fact <- read_csv("./data/validation-EGLL-2019.csv.gz", show_col_types = FALSE) |> 
  mutate(across(.cols = contains("_TIME"), .fns = lubridate::dmy_hms))

egll_20190801 <- egll_fact |> filter(lubridate::date(MVT_TIME) == lubridate::date(ptn)) |> 
  mutate(SEEN_OSN = FLTID %in% at_egll$FLTID)
```

```{r}
egll_20190801 |> summarise(MVTS = n(), .by = PHASE) |> 
  ggplot() + 
  geom_col( aes(x = PHASE, y = MVTS)) +
  geom_text(aes(x = PHASE, y = MVTS, label = MVTS), vjust = 1.5, color = "white") +
  labs(title = "EGLL - movement count - 20190801", x = NULL, y = NULL)
```
